---
title: "MATH40005 Coursework Spring 2022"
author: "Leo JONES, CID: 02036984"
output: pdf_document
---
# Question 1

## Question 1(a)


```{r}
#reading data and creating breaks
d=read.table("spendingsurvey.txt",header=TRUE,sep=",")
labels= list("19 and younger","20-29","30-39","40-49","50-59","60 and over")
cat<-c(0,19,29,39,49,59,100)
#sorting entries into age ranges
age=as.numeric(t(d)[2,])
d$ageranges <- cut(age, breaks = cat)
#summing the visits over the different ranges and then plotting
ag=aggregate(d$visits,by =list(d$ageranges),FUN=sum)
barplot(as.vector(ag$x), names= labels, cex.names = 0.75, col=terrain.colors(6),main="Visits by age group")
```


\clearpage

## Question 1(b)
```{r}
#same data as 1(a) but as a pie chart
pie(as.vector(ag$x), labels= labels, col=terrain.colors(6),main="Visits by age group")
```



\clearpage

## Question 1(c)
```{r}
#taking the mean of the total spend per visit for each age range
q1c=aggregate(d$totalspend/d$visits,by= list(d$ageranges), FUN=mean)
#plotting
barplot(as.vector(q1c$x), names =labels,cex.names=0.75,col=terrain.colors(6),main="Average spend (£) per visit by age group")
```

\clearpage

## Question 1(d)

The 30-39 group has the largest proportion of visits however the 50-59 group has the highest average spend.

```{r}
#taking the mean of the total spend for each age group
q1d=aggregate(d$totalspend,by= list(d$ageranges), FUN=mean)
print(q1d)
```

If you look at the total spend by age group, however, 30-39 has the highest spend, and therefore I would say they are most valuable to the restaurant chain.

\clearpage

## Question 1(e)
```{r}
#creating lists of the spends per visit for each age range
h=aggregate(d$totalspend/d$visits, by =list(d$ageranges), FUN=list)
#t-test without equality of variances
t.test(unlist(h$x[2]),unlist(h$x[4]))
```
Assuming normality but not equality of variances, the p-value is less than 0.05, which indicates a significant difference. 

We can use the F test to test for equality in the variances:

```{r}
#F-test to compare two variances
k
